\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{array}
\usepackage{geometry}
\usepackage{longtable}
\geometry{margin=1in}

\title{Overview}
\author{}
\date{}

\begin{document}

\maketitle

\section*{1. Language Design}

Hanami provides a one-to-one mapping between its “garden” vocabulary and canonical C++ constructs—e.g. \texttt{garden} $\rightarrow$ \texttt{namespace}, \texttt{species} $\rightarrow$ \texttt{class}, \texttt{bloom}/\texttt{water} $\rightarrow$ stream operators. Because all native C++ types, operators, and literals remain valid, users can freely intermix Hanami keywords with regular C++ expressions without extra lowering logic. The shared \texttt{TokenType} enumeration ensures every compiler phase agrees on the same keyword set, preventing drift between stages.

\section*{2. Compiler Pipeline}

\begin{longtable}{|>{\raggedright\arraybackslash}p{3.5cm}|>{\raggedright\arraybackslash}p{6.8cm}|>{\raggedright\arraybackslash}p{4.5cm}|}
\hline
\textbf{Stage} & \textbf{Implementation Highlights – What the Stage Does} & \textbf{Output} \\
\hline
Lexer – Lexical Analysis & 
• Reads the character stream once and groups into tokens (names, numbers, strings, operators, Hanami keywords). \newline
• Ignores single-line and multi-line comments; records exact line/column positions for later error reporting. \newline
• Detects and reports early errors like invalid characters, unclosed strings, etc. &
Token list with positions \\
\hline
Parser – Recursive-Descent Syntax Analysis &
• Receives tokens and builds an Abstract Syntax Tree (AST) representing program structure (gardens, species, functions, statements). \newline
• Uses manual recursive functions, supports panic-mode recovery: on syntax error, skips to synchronization point to continue catching more errors instead of stopping. &
Serializable AST (JSON) \\
\hline
Semantic Analysis + IR &
• Traverses AST to build symbol tables, apply scoping and visibility rules (open/hidden/guarded), perform static type checks (e.g. disallow \texttt{blossom} in \texttt{void} functions). \newline
• Labels each AST node with type and ID, emits language-neutral Intermediate Representation (JSON). &
Rich-context IR (JSON) \\
\hline
Code Generation / Transpiler &
• Consumes the unified IR and, via pluggable visitors, generates source code in C++, Java, Python, and JavaScript. \newline
• Ensures independent target generation: failure in one target does not affect others. \newline
• Retains original line numbers so runtime errors in generated code can be traced back to Hanami source. &
Files: \texttt{.cpp}, \texttt{.java}, \texttt{.py}, \texttt{.js} \\
\hline
\end{longtable}

\end{document}
